---
title: "humidity-cvd"
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: loaddata
#| echo: false

combined_data_relhum_forpanel <- readRDS(here::here("data","combined_data_relhum_forpanel_revised.rds"))
library(tidyverse)
gbd_data_clean <- read_csv(here::here("data","gbd_updated_april24.csv"))
gbd_data_clean <- gbd_data_clean |> 
        filter(Year %in% 2011:2019)

# rename state.name as State
combined_data_relhum_forpanel <- combined_data_relhum_forpanel |> 
 dplyr::rename(State = state_name,
              Year = year)

# merge humidity data and gbd data

combined_data <- merge(
  combined_data_relhum_forpanel,
  gbd_data_clean,
  by = c("State", "Year")
)



```

```{r}
#| label: libraries
#| echo: false
#| results: hide
#| message: false
#| warning: false
## data viz for humidity yerawise

pacman::p_load("tidyverse","ggcorrplot","directlabels","pglm", "scales")

```

### Correlation Analysis

```{r}
#| label: correlation
#| echo: false
#| results: hide
# Correlation analysis
combined_data <- combined_data |> 
        rename(maximum_25=high_25,
               maximum_975=high_975,
               minimum_25=low_25,
               minimum_975=low_975,
               average_25=avg_25,
               average_975=avg_975,
               median_25=med_25,
               median_975=med_975,) 
correlation_matrix <- cor(combined_data |> select(average_25, average_975, median_25,median_975,mode_25,mode_975,
                                                  maximum_25,maximum_975,
                                                  minimum_25,minimum_975,variation_25,variation_975), method = "pearson")
#correlation_matrix
# visualize as heatmap


# Compute a matrix of correlation p-values
p.mat <- cor_pmat(correlation_matrix)

```

```{r}
#| label: fig-corrplot
#| fig-cap: Correlation matrix for humidity variables
#| echo: false

ggcorrplot(correlation_matrix, 
           outline.color = "white",
           type = "lower",
           ggtheme = ggplot2::theme_gray,
           colors = c("#6D9EC1", "white", "#E46726"),
           lab = TRUE,
           p.mat = p.mat,
           insig = "blank")
```

```{r}
#| label: fig-relhumsummaries
#| echo: false
#| layout-ncol: 1
#| fig-cap: humidity summaries for India, 2011-2019
#| fig-subcap: 
#|     - "humidity Averages"
#|     - "humidity Variations"
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top

# averages
combined_data |>
        ggplot(aes(x = factor(Year), group=1)) +
        geom_line(aes(y=average_25),linetype="dashed", color="midnightblue")+
        geom_line(aes(y=average_975),linetype="dashed", color="midnightblue")+
        geom_ribbon(aes(ymin=average_25, ymax=average_975),fill="midnightblue", alpha=0.4)+
        facet_wrap(~State)+
        theme_minimal(base_size = 20)+
        labs(y="humidity (°C)", x="Year",
             title = "Average humidity",
             subtitle = "2.5th and 97.5th percentiles")+
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12))

# variations
combined_data |>
        ggplot(aes(x = factor(Year), group=1)) +
        geom_line(aes(y=variation_25),linetype="dashed", color="seagreen")+
        geom_line(aes(y=variation_975),linetype="dashed", color="seagreen")+
        geom_ribbon(aes(ymin=variation_25, ymax=variation_975),fill="seagreen", alpha=0.4)+
     #geom_ribbon(aes(ymin=0, ymax=variation_25),fill="seagreen", alpha=0.4)+
        facet_wrap(~State)+
        theme_minimal(base_size = 20)+
        labs(y="humidity (°C)", x="Year",
             title = "humidity Variations",
             subtitle = "2.5th percentiles")+
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12))


```

```{r}
#| label: panelregression
#| results: hide
#| echo: false
#| warning: false
#| message: false
model_CVD_Death_avg25 <- pglm(CVD_Death ~ average_25,
                              data = combined_data,
                              family = poisson,
                              model = "within",
                              index = c("State", "Year"))
summary(model_CVD_Death_avg25)
confint(model_CVD_Death_avg25)[, 1]
confint(model_CVD_Death_avg25)[, 2]
irr_avg25_death <- exp(coef(model_CVD_Death_avg25))
irr_avg25_death
irr_lci_avg25_death <- exp(confint(model_CVD_Death_avg25)[, 1])
irr_uci_avg25_death <- exp(confint(model_CVD_Death_avg25)[, 2])
irr_lci_avg25_death
irr_uci_avg25_death

model_CVD_Death_avg975 <- pglm(CVD_Death ~ average_975,
                              data = combined_data,
                              family = poisson,
                              model = "within",
                              index = c("State", "Year"))
summary(model_CVD_Death_avg975)
confint(model_CVD_Death_avg975)[, 1]
confint(model_CVD_Death_avg975)[, 2]
irr_avg975_death <- exp(coef(model_CVD_Death_avg975))
irr_avg975_death
irr_lci_avg975_death <- exp(confint(model_CVD_Death_avg975)[, 1])
irr_uci_avg975_death <- exp(confint(model_CVD_Death_avg975)[, 2])
irr_lci_avg975_death
irr_uci_avg975_death

model_CVD_Death_var25 <- pglm(CVD_Death ~ variation_25,
                               data = combined_data,
                               family = poisson,
                               model = "within",
                               index = c("State", "Year"))
summary(model_CVD_Death_var25)
irr_var25_death <- exp(coef(model_CVD_Death_var25))
irr_var25_death
irr_lci_var25_death <- exp(confint(model_CVD_Death_var25)[, 1])
irr_uci_var25_death <- exp(confint(model_CVD_Death_var25)[, 2])
irr_lci_var25_death
irr_uci_var25_death


### CVD_Incidence

model_CVD_Incidence_avg25 <- pglm(CVD_Incidence ~ average_25,
                                  data = combined_data,
                                  family = poisson,
                                  model = "within",
                                  index = c("State", "Year"))
summary(model_CVD_Incidence_avg25)
irr_avg25_incidence <- exp(coef(model_CVD_Incidence_avg25))
irr_avg25_incidence
irr_lci_avg25_incidence <- exp(confint(model_CVD_Incidence_avg25)[, 1])
irr_uci_avg25_incidence <- exp(confint(model_CVD_Incidence_avg25)[, 2])
irr_lci_avg25_incidence
irr_uci_avg25_incidence



model_CVD_Incidence_avg975 <- pglm(CVD_Incidence ~ average_975,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_Incidence_avg975)
irr_avg975_incidence <- exp(coef(model_CVD_Incidence_avg975))
irr_avg975_incidence
irr_lci_avg975_incidence <- exp(confint(model_CVD_Incidence_avg975)[, 1])
irr_uci_avg975_incidence <- exp(confint(model_CVD_Incidence_avg975)[, 2])
irr_lci_avg975_incidence
irr_uci_avg975_incidence


model_CVD_Incidence_var25 <- pglm(CVD_Incidence ~ variation_25,
                                  data = combined_data,
                                  family = poisson,
                                  model = "within",
                                  index = c("State", "Year"))
summary(model_CVD_Incidence_var25)
irr_var25_incidence <- exp(coef(model_CVD_Incidence_var25))
irr_var25_incidence
irr_lci_var25_incidence <- exp(confint(model_CVD_Incidence_var25)[, 1])
irr_uci_var25_incidence <- exp(confint(model_CVD_Incidence_var25)[, 2])
irr_lci_var25_incidence
irr_uci_var25_incidence


### CVD_Prevalence

model_CVD_Prevalence_avg25 <- pglm(CVD_Prevalence ~ average_25,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_Prevalence_avg25)
irr_avg25_prevalence <- exp(coef(model_CVD_Prevalence_avg25))
irr_avg25_prevalence
irr_lci_avg25_prevalence <- exp(confint(model_CVD_Prevalence_avg25)[, 1])
irr_uci_avg25_prevalence <- exp(confint(model_CVD_Prevalence_avg25)[, 2])
irr_lci_avg25_prevalence
irr_uci_avg25_prevalence



model_CVD_Prevalence_avg975 <- pglm(CVD_Prevalence ~ average_975,
                                    data = combined_data,
                                    family = poisson,
                                    model = "within",
                                    index = c("State", "Year"))
summary(model_CVD_Prevalence_avg975)
irr_avg975_prevalence <- exp(coef(model_CVD_Prevalence_avg975))
irr_avg975_prevalence
irr_lci_avg975_prevalence <- exp(confint(model_CVD_Prevalence_avg975)[, 1])
irr_uci_avg975_prevalence <- exp(confint(model_CVD_Prevalence_avg975)[, 2])
irr_lci_avg975_prevalence
irr_uci_avg975_prevalence


model_CVD_Prevalence_var25 <- pglm(CVD_Prevalence ~ variation_25,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_Prevalence_var25)
irr_var25_prevalence <- exp(coef(model_CVD_Prevalence_var25))
irr_var25_prevalence
irr_lci_var25_prevalence <- exp(confint(model_CVD_Prevalence_var25)[, 1])
irr_uci_var25_prevalence <- exp(confint(model_CVD_Prevalence_var25)[, 2])
irr_lci_var25_prevalence
irr_uci_var25_prevalence


```

### IRR

| Humidity | Percentile | Mortality \[95% CIs\]                                                                                     | Prevalence\[95% CIs\]                                                                                                   | Incidence\[95% CIs\]                                                                                                 |
|---------------|---------------|---------------|---------------|---------------|
| Average           | 2.5        | **`r round(irr_avg25_death,3)` \[`r round(irr_lci_avg25_death,3)`,`r round(irr_uci_avg25_death,3)`\]**    | **`r round(irr_avg25_prevalence,3)`\[`r round(irr_lci_avg25_prevalence,3)`,`r round(irr_uci_avg25_prevalence,3)`\]**    | **`r round(irr_avg25_incidence,3)`\[`r round(irr_lci_avg25_incidence,3)`,`r round(irr_uci_avg25_incidence,3)`\]**    |
| Average           | 97.5       | **`r round(irr_avg975_death,3)` \[`r round(irr_lci_avg975_death,3)`,`r round(irr_uci_avg975_death,3)`\]** | **`r round(irr_avg975_prevalence,3)`\[`r round(irr_lci_avg975_prevalence,3)`,`r round(irr_uci_avg975_prevalence,3)`\]** | **`r round(irr_avg975_incidence,3)`\[`r round(irr_lci_avg975_incidence,3)`,`r round(irr_uci_avg975_incidence,3)`\]** |
| Variation         | 2.5        | `r round(irr_var25_death,3)` \[`r round(irr_lci_var25_death,3)`,`r round(irr_uci_var25_death,3)`\]        | **`r round(irr_var25_prevalence,3)`\[`r round(irr_lci_var25_prevalence,3)`,`r round(irr_uci_var25_prevalence,3)`\]**    | **`r round(irr_var25_incidence,3)`\[`r round(irr_lci_var25_incidence,3)`,`r round(irr_uci_var25_incidence,3)`\]**    |

: Incidence Rate Ratios derived from Poisson panel regressions for CVD burden (mortality, prevalence, and incidence) {#tbl-my-table}


### PAF

```{r}
#| label: pafcalc
#| results: hide
#| echo: false
#| warning: false
#| message: false
gbd_data_uncorrected <- read_csv(here::here("data","gbd_data_uncorrected.csv"))
gbd_data_uncorrected <- gbd_data_uncorrected |> 
        filter(Year %in% 2011:2019)

# rename state.name as State
#combined_data_temperature_forpanel 
#<- combined_data_temperature_forpanel |> 
 #dplyr::rename(State = state_name,
 #             Year = year)

# merge temperature data and gbd data

combined_data_uncorrectedgbd <- merge(
  combined_data_relhum_forpanel,
  gbd_data_uncorrected,
  by = c("State", "Year")
)

paf_data_formapviz <- combined_data_uncorrectedgbd |>
dplyr::select(State, Year, CVD_Death, avg_975,
              CVD_Prevalence, CVD_Incidence) |>
filter(Year==2019) |> 
mutate(paf= round((CVD_Death * 0.002)),
       paf_prev= round(CVD_Prevalence * 0.002),
       paf_incid=round(CVD_Incidence * 0.002)) |> 
dplyr::select(-CVD_Death,-avg_975) |> 
        #rename State as state
        dplyr::rename(state = State)
       

```


```{r}
#| label: fig-pafmap
#| echo: false
#| layout-ncol: 1
#| fig-cap: CVD deaths attributable to high average temperature in Indian states
#| fig-subcap: 
#|     - CVD Mortality
#|     - CVD Incidence
#|     - CVD Prevalence
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top
#| message: false
#| warning: false
library(sf)
library(tidyverse)

india <- readRDS(here::here("data",
                            "spatial_files", 
                            "India_states.rds"))
india <- india |> 
        rename(state = NAME_1) |> 
        # recode orissa as odisha
        mutate(state = recode(state, "Orissa" = "Odisha")) |> 
# recode uttaranchal as uttarakhand
        mutate(state = recode(state, "Uttaranchal" = "Uttarakhand"))


# combine paf_data_formapviz with india based on state variable
india_paf <- india |> 
  left_join(paf_data_formapviz, by = "state")
india_paf <- india_paf |> 
        mutate(centroid = st_centroid(geometry)) %>% 
  mutate(x = st_coordinates(centroid)[,1], 
         y = st_coordinates(centroid)[,2])


# mortality map
india_paf |> 
  ggplot() +
  geom_sf(aes(fill = paf)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf), size = 3, color = "white") +
  labs(fill = "CVD deaths",
       title = "CVD deaths attributable to high average humidity in Indian states",
       subtitle = "Year 2019") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())

# Incidence map
india_paf |> 
  ggplot() +
  geom_sf(aes(fill = paf_incid)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf_incid), size = 3, color = "white") +
  labs(fill = "CVD Incidence",
       title = "CVD Incidence attributable to high average humidity in Indian states",
       subtitle = "Year 2019") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())


# Prevalence map
india_paf |> 
  ggplot() +
  geom_sf(aes(fill = paf_prev)) +
  scale_fill_gradient(low = "orange", high = "midnightblue",
                      labels=scales::comma) +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf_prev), size = 3, color = "white") +
  labs(fill = "CVD Prevalence",
       title = "CVD Prevalence attributable to high average humidity in Indian states",
       subtitle = "Year 2019") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())

```

