---
title: "Humidity-CVD"
format:
  html:
    code-fold: true
    code-summary: "Click here to view/hide the code"
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Data loading

```{r}
#| label: loaddata
#| echo: false
#| warning: false

combined_data_relhum_forpanel <- readRDS(here::here("data","combined_data_relhum_forpanel_revised.rds"))
suppressPackageStartupMessages(library(tidyverse))

#read excel file

library(readxl)

data <- read_excel(here::here(
  "data", "CVD_DALY_YLD_YLL_1990-2021_per100k.xlsx"
))

library(tidyverse)
data <- data |> 
  filter(Year %in% 2011:2022) |> 
  dplyr::rename(State=Location)

#gbd_data_clean <- read_csv(here::here("data","gbd_updated_april24.csv"), show_col_types = FALSE)
#gbd_data_clean <- gbd_data_clean |> 
 #       filter(Year %in% 2011:2019)

# rename state.name as State
combined_data_relhum_forpanel <- combined_data_relhum_forpanel |> 
 dplyr::rename(State = state_name,
              Year = year)

# merge humidity data and gbd data

combined_data <- merge(
  combined_data_relhum_forpanel,
  data,
  by = c("State", "Year")
)



```

## Packages

```{r}
#| label: libraries
#| echo: true
#| results: hide
#| message: false
#| warning: false
## data viz for humidity yerawise

pacman::p_load("ggcorrplot","directlabels","pglm", "scales")
suppressPackageStartupMessages(library(tidyverse))


```

------------------------------------------------------------------------

### Correlation Analysis

```{r}
#| label: correlation
#| echo: true
#| results: hide
# Correlation analysis
combined_data <- combined_data |> 
        rename(maximum_25=high_25,
               maximum_975=high_975,
               minimum_25=low_25,
               minimum_975=low_975,
               average_25=avg_25,
               average_975=avg_975,
               median_25=med_25,
               median_975=med_975,) 
correlation_matrix <- cor(combined_data |> select(average_25, average_975, median_25,median_975,mode_25,mode_975,
                                                  maximum_25,maximum_975,
                                                  minimum_25,minimum_975,variation_25,variation_975), method = "pearson")
#correlation_matrix
# visualize as heatmap


# Compute a matrix of correlation p-values
p.mat <- cor_pmat(correlation_matrix)

```

```{r}
#| label: fig-corrplot
#| fig-cap: Correlation matrix for humidity variables
#| echo: true
#| column: page

ggcorrplot(correlation_matrix, 
           outline.color = "white",
           type = "lower",
           ggtheme = ggplot2::theme_gray,
           colors = c("#6D9EC1", "white", "#E46726"),
           lab = TRUE,
           p.mat = p.mat,
           insig = "blank")
```

*This matrix shows pairwise correlations among various humidity metrics (average, median, mode, minimum, maximum, variation) at the 2.5th and 97.5th percentiles. Only statistically significant correlations are displayed, indicating strong associations among several metrics. Based on this analysis, both 2.5 and 97.5 percentiles of the average humidity were selected for further analysis. This is to minimize multicollinearity and retain representative variability in the downstream panel regression.*

------------------------------------------------------------------------

```{r}
#| label: fig-relhumsummaries
#| echo: true
#| layout-ncol: 1
#| fig-cap: Humidity summaries for India, 2011-2019
#| fig-subcap: 
#|     - "Average humidity levels (2.5th and 97.5th percentiles) across Indian states from 2011 to 2019."
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top
#| column: screen-inset-shaded

# averages
combined_data |>
        ggplot(aes(x = factor(Year), group=1)) +
        geom_line(aes(y=average_25),linetype="dashed", color="midnightblue")+
        geom_line(aes(y=average_975),linetype="dashed", color="midnightblue")+
        geom_ribbon(aes(ymin=average_25, ymax=average_975),fill="midnightblue", alpha=0.4)+
        facet_wrap(~State)+
        theme_minimal(base_size = 20)+
        labs(y="humidity (%)", x="Year",
             title = "Average humidity",
             subtitle = "2.5th and 97.5th percentiles")+
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12))



```

@fig-relhumsummaries-1 illustrates the trends in the 2.5th and 97.5th percentiles of daily humidity for each year across Indian states from 2011 to 2019. The 2.5th percentile of annual daily humidity remains consistently below 25% across all states. The lowest value was observed in Gujarat in 2011 (13.8%), while the highest value at this percentile was recorded in Tamil Nadu in 2019 (22.1%). This indicates a limited range of variation at the lower end of the humidity spectrum across states. The 97.5th percentile of annual daily humidity showed a wider range of variability across states, with the lowest value recorded in Odisha in 2011 (37.3%) and the highest value in Uttarakhand in 2019 (87.2%). While most states consistently recorded values below 75%, certain states crossed this threshold in specific years. For instance, Uttarakhand, Sikkim, Rajasthan, Punjab, Haryana, Himachal Pradesh, Delhi, Bihar, and Arunachal Pradesh recorded 97.5th percentile values above 75% in 2019. Similarly, northeastern states like Nagaland, Meghalaya, and Manipur crossed this threshold in 2013. These observations indicate episodic peaks in maximum humidity levels in certain states, particularly during the later years of the study period.



------------------------------------------------------------------------

```{r}
#| label: panelregression
#| results: hide
#| echo: true
#| warning: false
#| message: false
model_CVD_Death_avg25 <- pglm(CVD_Death ~ average_25,
                              data = combined_data,
                              family = poisson,
                              model = "within",
                              index = c("State", "Year"))
summary(model_CVD_Death_avg25)
confint(model_CVD_Death_avg25)[, 1]
confint(model_CVD_Death_avg25)[, 2]
irr_avg25_death <- exp(coef(model_CVD_Death_avg25))
irr_avg25_death
irr_lci_avg25_death <- exp(confint(model_CVD_Death_avg25)[, 1])
irr_uci_avg25_death <- exp(confint(model_CVD_Death_avg25)[, 2])
irr_lci_avg25_death
irr_uci_avg25_death

model_CVD_Death_avg975 <- pglm(CVD_Death ~ average_975,
                              data = combined_data,
                              family = poisson,
                              model = "within",
                              index = c("State", "Year"))
summary(model_CVD_Death_avg975)
confint(model_CVD_Death_avg975)[, 1]
confint(model_CVD_Death_avg975)[, 2]
irr_avg975_death <- exp(coef(model_CVD_Death_avg975))
irr_avg975_death
irr_lci_avg975_death <- exp(confint(model_CVD_Death_avg975)[, 1])
irr_uci_avg975_death <- exp(confint(model_CVD_Death_avg975)[, 2])
irr_lci_avg975_death
irr_uci_avg975_death


### CVD_Incidence

model_CVD_Incidence_avg25 <- pglm(CVD_Incidence ~ average_25,
                                  data = combined_data,
                                  family = poisson,
                                  model = "within",
                                  index = c("State", "Year"))
summary(model_CVD_Incidence_avg25)
irr_avg25_incidence <- exp(coef(model_CVD_Incidence_avg25))
irr_avg25_incidence
irr_lci_avg25_incidence <- exp(confint(model_CVD_Incidence_avg25)[, 1])
irr_uci_avg25_incidence <- exp(confint(model_CVD_Incidence_avg25)[, 2])
irr_lci_avg25_incidence
irr_uci_avg25_incidence



model_CVD_Incidence_avg975 <- pglm(CVD_Incidence ~ average_975,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_Incidence_avg975)
irr_avg975_incidence <- exp(coef(model_CVD_Incidence_avg975))
irr_avg975_incidence
irr_lci_avg975_incidence <- exp(confint(model_CVD_Incidence_avg975)[, 1])
irr_uci_avg975_incidence <- exp(confint(model_CVD_Incidence_avg975)[, 2])
irr_lci_avg975_incidence
irr_uci_avg975_incidence



### CVD_Prevalence

model_CVD_Prevalence_avg25 <- pglm(CVD_Prevalence ~ average_25,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_Prevalence_avg25)
irr_avg25_prevalence <- exp(coef(model_CVD_Prevalence_avg25))
irr_avg25_prevalence
irr_lci_avg25_prevalence <- exp(confint(model_CVD_Prevalence_avg25)[, 1])
irr_uci_avg25_prevalence <- exp(confint(model_CVD_Prevalence_avg25)[, 2])
irr_lci_avg25_prevalence
irr_uci_avg25_prevalence



model_CVD_Prevalence_avg975 <- pglm(CVD_Prevalence ~ average_975,
                                    data = combined_data,
                                    family = poisson,
                                    model = "within",
                                    index = c("State", "Year"))
summary(model_CVD_Prevalence_avg975)
irr_avg975_prevalence <- exp(coef(model_CVD_Prevalence_avg975))
irr_avg975_prevalence
irr_lci_avg975_prevalence <- exp(confint(model_CVD_Prevalence_avg975)[, 1])
irr_uci_avg975_prevalence <- exp(confint(model_CVD_Prevalence_avg975)[, 2])
irr_lci_avg975_prevalence
irr_uci_avg975_prevalence




```

### IRR - cvd prevalence incidence mortality


| Humidity  | Percentile | Mortality \[95% CIs\]                                                                                     | Prevalence\[95% CIs\]                                                                                                   | Incidence\[95% CIs\]                                                                                                 |
|-----------|------------|-----------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------|
| Average   | 2.5        | **`r round(irr_avg25_death,3)` \[`r round(irr_lci_avg25_death,3)`,`r round(irr_uci_avg25_death,3)`\]**    | **`r round(irr_avg25_prevalence,3)`\[`r round(irr_lci_avg25_prevalence,3)`,`r round(irr_uci_avg25_prevalence,3)`\]**    | **`r round(irr_avg25_incidence,3)`\[`r round(irr_lci_avg25_incidence,3)`,`r round(irr_uci_avg25_incidence,3)`\]**    |
| Average   | 97.5       | **`r round(irr_avg975_death,3)` \[`r round(irr_lci_avg975_death,3)`,`r round(irr_uci_avg975_death,3)`\]** | **`r round(irr_avg975_prevalence,3)`\[`r round(irr_lci_avg975_prevalence,3)`,`r round(irr_uci_avg975_prevalence,3)`\]** | **`r round(irr_avg975_incidence,3)`\[`r round(irr_lci_avg975_incidence,3)`,`r round(irr_uci_avg975_incidence,3)`\]** |


: Incidence Rate Ratios derived from Poisson panel regressions for CVD burden (mortality, prevalence, and incidence) {#tbl-my-table}

------------------------------------------------------------------------

### PAF

```{r}
#| label: pafcalc
#| results: hide
#| echo: true
#| warning: false
#| message: false
gbd_data_uncorrected_new <- read_xlsx(here::here("data","CVD_DALY_YLD_YLL_1990-2021_per100k.xlsx"))
gbd_data_uncorrected_new <- gbd_data_uncorrected_new |> 
        filter(Year %in% 2011:2021)

gbd_data_uncorrected_new <- gbd_data_uncorrected_new |> 
  dplyr::rename(State=Location)

# rename state.name as State
#combined_data_temperature_forpanel 
#<- combined_data_temperature_forpanel |> 
 #dplyr::rename(State = state_name,
 #             Year = year)

# merge temperature data and gbd data

combined_data_uncorrectedgbd <- merge(
  combined_data_relhum_forpanel,
  gbd_data_uncorrected_new,
  by = c("State", "Year")
)

paf_data_formapviz <- combined_data_uncorrectedgbd |>
dplyr::select(State, Year, CVD_Death, avg_975,
              CVD_Prevalence, CVD_Incidence) |>
dplyr::filter(Year==2019) |> 
mutate(paf= round((CVD_Death * 0.002)),
       paf_prev= round(CVD_Prevalence * 0.002),
       paf_incid=round(CVD_Incidence * 0.002)) |> 
dplyr::select(-CVD_Death,-avg_975) |> 
        #rename State as state
        dplyr::rename(state = State)
       

```

```{r}
#| label: fig-pafmap
#| echo: true
#| layout-ncol: 1
#| fig-cap: CVD deaths attributable to high average humidity in Indian states
#| fig-subcap: 
#|     - CVD Mortality
#|     - CVD Incidence
#|     - CVD Prevalence
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top
#| message: false
#| warning: false
#| column: screen
#| out-width: 100%
library(sf)
library(tidyverse)

india <- readRDS(here::here("data",
                            "spatial_files", 
                            "India_states.rds"))
india <- india |> 
        rename(state = NAME_1) |> 
        # recode orissa as odisha
        mutate(state = recode(state, "Orissa" = "Odisha")) |> 
# recode uttaranchal as uttarakhand
        mutate(state = recode(state, "Uttaranchal" = "Uttarakhand"))


# combine paf_data_formapviz with india based on state variable
india_paf <- india |> 
  left_join(paf_data_formapviz, by = "state")
india_paf <- india_paf |> 
        mutate(centroid = st_centroid(geometry)) %>% 
  mutate(x = st_coordinates(centroid)[,1], 
         y = st_coordinates(centroid)[,2])


# mortality map
india_paf |> 
  ggplot() +
  geom_sf(aes(fill = paf)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf), size = 3, color = "white") +
  labs(fill = "CVD deaths",
       title = "CVD deaths attributable to high average humidity in Indian states",
       subtitle = "Year 2019") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())

# Incidence map
india_paf |> 
  ggplot() +
  geom_sf(aes(fill = paf_incid)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf_incid), size = 3, color = "white") +
  labs(fill = "CVD Incidence",
       title = "CVD Incidence attributable to high average humidity in Indian states",
       subtitle = "Year 2019") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())


# Prevalence map
india_paf |> 
  ggplot() +
  geom_sf(aes(fill = paf_prev)) +
  scale_fill_gradient(low = "orange", high = "midnightblue",
                      labels=scales::comma) +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf_prev), size = 3, color = "white") +
  labs(fill = "CVD Prevalence",
       title = "CVD Prevalence attributable to high average humidity in Indian states",
       subtitle = "Year 2019") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())

```

------------------------------------------------------------------------

```{r}
#| label: panelreg-economy
#| results: hide
#| echo: true
#| warning: false
#| message: false

## daly

model_CVD_DALY_avg25 <- pglm(CVD_DALY ~ average_25,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_DALY_avg25)
irr_avg25_DALY<- exp(coef(model_CVD_DALY_avg25))
irr_avg25_DALY
irr_lci_avg25_DALY <- exp(confint(model_CVD_DALY_avg25)[, 1])
irr_uci_avg25_DALY <- exp(confint(model_CVD_DALY_avg25)[, 2])
irr_lci_avg25_DALY
irr_uci_avg25_DALY



model_CVD_DALY_avg975 <- pglm(CVD_DALY ~ average_975,
                                    data = combined_data,
                                    family = poisson,
                                    model = "within",
                                    index = c("State", "Year"))
summary(model_CVD_DALY_avg975)
irr_avg975_DALY <- exp(coef(model_CVD_DALY_avg975))
irr_avg975_DALY
irr_lci_avg975_DALY  <- exp(confint(model_CVD_DALY_avg975)[, 1])
irr_uci_avg975_DALY  <- exp(confint(model_CVD_DALY_avg975)[, 2])
irr_lci_avg975_DALY 
irr_uci_avg975_DALY 



## yld

model_CVD_YLD_avg25 <- pglm(CVD_YLD ~ average_25,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_YLD_avg25)
irr_avg25_YLD <- exp(coef(model_CVD_YLD_avg25))
irr_avg25_YLD
irr_lci_avg25_YLD <- exp(confint(model_CVD_YLD_avg25)[, 1])
irr_uci_avg25_YLD <- exp(confint(model_CVD_YLD_avg25)[, 2])
irr_lci_avg25_YLD
irr_uci_avg25_YLD



model_CVD_YLD_avg975 <- pglm(CVD_YLD ~ average_975,
                                    data = combined_data,
                                    family = poisson,
                                    model = "within",
                                    index = c("State", "Year"))
summary(model_CVD_YLD_avg975)
irr_avg975_YLD <- exp(coef(model_CVD_YLD_avg975))
irr_avg975_YLD
irr_lci_avg975_YLD  <- exp(confint(model_CVD_YLD_avg975)[, 1])
irr_uci_avg975_YLD  <- exp(confint(model_CVD_YLD_avg975)[, 2])
irr_lci_avg975_YLD 
irr_uci_avg975_YLD 



## YLL

model_CVD_YLL_avg25 <- pglm(CVD_YLL ~ average_25,
                                   data = combined_data,
                                   family = poisson,
                                   model = "within",
                                   index = c("State", "Year"))
summary(model_CVD_YLL_avg25)
irr_avg25_YLL <- exp(coef(model_CVD_YLL_avg25))
irr_avg25_YLL
irr_lci_avg25_YLL <- exp(confint(model_CVD_YLL_avg25)[, 1])
irr_uci_avg25_YLL <- exp(confint(model_CVD_YLL_avg25)[, 2])
irr_lci_avg25_YLL
irr_uci_avg25_YLL



model_CVD_YLL_avg975 <- pglm(CVD_YLL ~ average_975,
                                    data = combined_data,
                                    family = poisson,
                                    model = "within",
                                    index = c("State", "Year"))
summary(model_CVD_YLL_avg975)
irr_avg975_YLL <- exp(coef(model_CVD_YLL_avg975))
irr_avg975_YLL
irr_lci_avg975_YLL  <- exp(confint(model_CVD_YLL_avg975)[, 1])
irr_uci_avg975_YLL  <- exp(confint(model_CVD_YLL_avg975)[, 2])
irr_lci_avg975_YLL 
irr_uci_avg975_YLL 



```

### IRR - cvd DALY YLD YLL

| Humidity  | Percentile | DALY \[95% CIs\]                                                                                       | YLD\[95% CIs\]                                                                                     | YLL\[95% CIs\]                                                                                     |
|-----------|------------|--------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------|
| Average   | 2.5        | **`r round(irr_avg25_DALY,3)` \[`r round(irr_lci_avg25_DALY,3)`,`r round(irr_uci_avg25_DALY,3)`\]**    | **`r round(irr_avg25_YLD,3)`\[`r round(irr_lci_avg25_YLD,3)`,`r round(irr_uci_avg25_YLD,3)`\]**    | `r round(irr_avg25_YLL,3)`\[`r round(irr_lci_avg25_YLL,3)`,`r round(irr_uci_avg25_YLL,3)`\]        |
| Average   | 97.5       | **`r round(irr_avg975_DALY,3)` \[`r round(irr_lci_avg975_DALY,3)`,`r round(irr_uci_avg975_DALY,3)`\]** | **`r round(irr_avg975_YLD,3)`\[`r round(irr_lci_avg975_YLD,3)`,`r round(irr_uci_avg975_YLD,3)`\]** | **`r round(irr_avg975_YLL,3)`\[`r round(irr_lci_avg975_YLL,3)`,`r round(irr_uci_avg975_YLL,3)`\]** |


: Incidence Rate Ratios derived from Poisson panel regressions for CVD burden (DALY, YLD, and YLL) {#tbl-my-table2}

### PAF for DALY, YLL, YLD

```{r}
#| label: pafcalc-economy
#| results: hide
#| echo: true
#| warning: false
#| message: false

gbd_data_uncorrected_new <- read_xlsx(here::here("data","CVD_DALY_YLL_YLD_2011-2019_Statewise_Uncorrected.xlsx")) |> 
  janitor::clean_names()
gbd_data_uncorrected_new <- gbd_data_uncorrected_new |> 
  dplyr::rename(State=location_2,
                Year=year_7,
                CVD_DALY=daly_val,
                CVD_YLD=yld_val,
                CVD_YLL=yll_val) |> 
        filter(Year %in% 2011:2021) |> 
  dplyr::select(Year,State,CVD_DALY,CVD_YLL,CVD_YLD)


# rename state.name as State
#combined_data_temperature_forpanel 
#<- combined_data_temperature_forpanel |> 
 #dplyr::rename(State = state_name,
 #             Year = year)

# merge temperature data and gbd data

gbd_data_uncorrected_economy_forpaf <- merge(
  combined_data_relhum_forpanel,
  gbd_data_uncorrected_new,
  by = c("State", "Year")
)

paf_economydata_formapviz <- gbd_data_uncorrected_economy_forpaf |>
dplyr::select(State, Year,  avg_975,
             CVD_DALY,CVD_YLL,CVD_YLD) |>
filter(Year==2021) |> 
mutate(paf_daly= round((CVD_DALY * 0.001)),
       paf_yld= round(CVD_YLL * 0.001),
       paf_yll=round(CVD_YLD * 0.001)) |> 
#dplyr::select(-CVD_Death,-avg_975) |> 
        #rename State as state
        dplyr::rename(state = State)
       
paf_rh <- gbd_data_uncorrected_economy_forpaf |>
dplyr::select(State, Year,  avg_975,
             CVD_DALY,CVD_YLL,CVD_YLD) |>
filter(Year==2022) |> 
mutate(paf_daly_avg_rh= round((CVD_DALY * 0.001))) |> 
#dplyr::select(-CVD_Death,-avg_975) |> 
        #rename State as state
        dplyr::rename(state = State)
write.csv(x = paf_rh,file = here::here(
        "output", "paf_values_rh.csv"
),row.names = FALSE)

write_rds(x = paf_rh,file = here::here(
        "output", "paf_values_rh.rds"))
```

```{r}
#| label: fig-pafmap-economy
#| echo: true
#| layout-ncol: 1
#| fig-cap: CVD burden attributable to high average humidity in Indian states
#| fig-subcap: 
#|     - CVD DALY
#|     - CVD YLD
#|     - CVD YLL
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top
#| message: false
#| warning: false
#| column: screen
#| out-width: 100%
library(sf)
library(tidyverse)

india <- readRDS(here::here("data",
                            "spatial_files", 
                            "India_states.rds"))
india <- india |> 
        rename(state = NAME_1) |> 
        # recode orissa as odisha
        mutate(state = recode(state, "Orissa" = "Odisha")) |> 
# recode uttaranchal as uttarakhand
        mutate(state = recode(state, "Uttaranchal" = "Uttarakhand"))


# combine paf_data_formapviz with india based on state variable
india_paf_economy <- india |> 
  left_join(paf_economydata_formapviz, by = "state")
india_paf_economy <- india_paf_economy |> 
        mutate(centroid = st_centroid(geometry)) %>% 
  mutate(x = st_coordinates(centroid)[,1], 
         y = st_coordinates(centroid)[,2])


# DALY map
india_paf_economy |> 
  ggplot() +
  geom_sf(aes(fill = paf_daly)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf_daly), size = 3, color = "white") +
  labs(fill = "CVD DALY",
       title = "CVD DALY attributable to high average humidity in Indian states",
       subtitle = "Year 2021") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())

# yld map
india_paf_economy |> 
  ggplot() +
  geom_sf(aes(fill = paf_yld)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf_yld), size = 3, color = "white") +
  labs(fill = "CVD YLD",
       title = "CVD YLD attributable to high average humidity in Indian states",
       subtitle = "Year 2021") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())


# YLL map
india_paf_economy |> 
  ggplot() +
  geom_sf(aes(fill = paf_yll)) +
  scale_fill_gradient(low = "orange", high = "midnightblue",
                      labels=scales::comma) +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf_yll), size = 3, color = "white") +
  labs(fill = "CVD YLL",
       title = "CVD YLL attributable to high average humidity in Indian states",
       subtitle = "Year 2021") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())

```

------------------------------------------------------------------------

## Projection maps for 97.5th percentile of average relative humidity

### CVD Deaths

```{r}
#| label: projmaps-deaths
#| warning: false
#| message: false

all_forecasts_deaths <- read_csv(here::here("output", "cvd_deaths","combined_forecasts_all_states.csv"))

all_forecasts_2030_deaths <- all_forecasts_deaths |>
  dplyr::filter(Year == 2030) |>
  mutate(paf= round((Forecasted_CVD_Deaths * 0.002),0)) |>
  dplyr::rename(state = State)

rmarkdown::paged_table(all_forecasts_2030_deaths)

```

```{r}
#| label: fig-projmaps-deaths
#| echo: true
#| layout-ncol: 1
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top
#| message: false
#| warning: false
#| column: screen
#| out-width: 100%
india <- readRDS(here::here("data",
                            "spatial_files",
                            "India_states.rds"))
india <- india |>
  rename(state = NAME_1) |>
  # recode orissa as odisha
  mutate(state = recode(state, "Orissa" = "Odisha")) |>
  # recode uttaranchal as uttarakhand
  mutate(state = recode(state, "Uttaranchal" = "Uttarakhand"))


# combine paf_data_formapviz with india based on state variable
india_paf_deaths <- india |>
  left_join(all_forecasts_2030_deaths, by = "state")
india_paf_deaths <- india_paf_deaths |>
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(x = st_coordinates(centroid)[,1],
         y = st_coordinates(centroid)[,2])



# mortality map
india_paf_deaths |>
  ggplot() +
  geom_sf(aes(fill = paf)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf), size = 3, color = "white") +
  labs(fill = "CVD deaths",
       title = "CVD deaths attributable to high average relative humidity in Indian states",
       subtitle = "Year 2030") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())

```

### CVD Prevalence

```{r}
#| label: projmaps-prev
#| warning: false
#| message: false


all_forecasts_prev <- read_csv(here::here("output", "cvd_prev","combined_forecasts_all_states.csv"))

all_forecasts_2030_prev <- all_forecasts_prev |>
  dplyr::filter(Year == 2030) |>
  mutate(paf= round((Forecasted_CVD_Prevalences * 0.002),0)) |>
  dplyr::rename(state = State)

rmarkdown::paged_table(all_forecasts_2030_prev)
```

```{r}
#| label: fig-projmaps-prev
#| echo: true
#| layout-ncol: 1
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top
#| message: false
#| warning: false
#| column: screen
#| out-width: 100%


# combine paf_data_formapviz with india based on state variable
india_paf_prev <- india |>
  left_join(all_forecasts_2030_prev, by = "state")
india_paf_prev <- india_paf_prev |>
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(x = st_coordinates(centroid)[,1],
         y = st_coordinates(centroid)[,2])



# prevalence map
india_paf_prev |>
  ggplot() +
  geom_sf(aes(fill = paf)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf), size = 3, color = "white") +
  labs(fill = "CVD prevalence",
       title = "CVD prevalence attributable to high average relative humidity in Indian states",
       subtitle = "Year 2030") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())





```

### CVD Incidence

```{r}
#| label: projmaps-inci
#| warning: false
#| message: false

all_forecasts_incidence <- read_csv(here::here("output", "cvd_inci","combined_forecasts_all_states.csv"))

all_forecasts_2030_incidence <- all_forecasts_incidence |>
  dplyr::filter(Year == 2030) |>
  mutate(paf= round((Forecasted_CVD_Incidences * 0.002),0)) |>
  dplyr::rename(state = State)

rmarkdown::paged_table(all_forecasts_2030_incidence)
```

```{r}
#| label: fig-projmaps-inci
#| echo: true
#| layout-ncol: 1
#| fig-height: 10
#| fig-width: 14
#| fig-align: center
#| fig-cap-location: top
#| message: false
#| warning: false
#| column: screen
#| out-width: 100%


# combine paf_data_formapviz with india based on state variable
india_paf_inci <- india |>
  left_join(all_forecasts_2030_incidence, by = "state")
india_paf_inci <- india_paf_inci |>
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(x = st_coordinates(centroid)[,1],
         y = st_coordinates(centroid)[,2])



# incidence map
india_paf_inci |>
  ggplot() +
  geom_sf(aes(fill = paf)) +
  scale_fill_gradient(low = "orange", high = "midnightblue") +
  # Add paf on the states as text with x and y coordinates
  geom_text(aes(x = x, y = y, label = paf), size = 3, color = "white") +
  labs(fill = "CVD incidence",
       title = "CVD incidence attributable to high average relative humidity in Indian states",
       subtitle = "Year 2030") +
  theme_classic() +
  # Make x-axis and y-axis blank
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank())



```


